---
title: "2016 Parties in New York"
author: "Tinguaro Barreno"
date: "November 3, 2017"
output:
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library-load, echo=FALSE, message=FALSE, warning=FALSE}
library(plotly)
```

## Introduction

The New York city hotline receives non-urgent community concerns, which are made public by
the city through [NYC Open Data](https://opendata.cityofnewyork.us/) portal.

The dataset that we've chosen focuses on Noise complaints that were collected in 2016 and
indicate ongoing party in a given neighborhood.

In this report we will show few different plots around this data set.

## Getting the data

I've chosen this data set beause it includes time series (creation/close date), grouped information (incidents by city/borough) and geoposition of the incidents.

The data has been obtained from the [Kaggle](https://www.kaggle.com/) dataset
'[partyntc](https://www.kaggle.com/somesnm/partynyc)'.

```{r getting-the-data, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
library( lubridate )

# Loading the data
parties <- read.csv( "party_in_nyc.csv" )

# Format conversion
parties$Created.Date <-ymd_hms( as.character(parties$Created.Date) )
parties$Closed.Date <- ymd_hms( as.character(parties$Closed.Date) )
```

## Data explanatory analysis

The data set contains 225,414 incidents (observations) with this structure:

```{r data-explanatory-analysis, echo=FALSE}
# Data set structure
str( parties )
```

## Incidents by month

This time series plot show the number of noise complains aggregated by month in 2016.

```{r plot-time-series, echo=FALSE, width="100%"}
# First: we need to count the incidents by month. As seen in StackOverflow:
# https://stackoverflow.com/questions/15181147/summarise-data-into-monthly-counts-by-year
parties_months <- data.frame( month=format( as.POSIXct(parties$Created.Date), "%Y-%m" ) )
incidents_by_month <- as.data.frame( table(parties_months) )

# Building the plot
plot_ly( x = ~incidents_by_month$parties_months,
         y = ~incidents_by_month$Freq,
         type = "scatter",
         mode = "lines+markers",
         marker = list( size = 15, color = "#52006d" ),
         showlegend = FALSE,
         hoverinfo = "text",
         text = paste0( "<br><b>", incidents_by_month$parties_months,
                        "</b></br> Incidents: ",
                        incidents_by_month$Freq) ) %>%
  layout( showlegend = FALSE,
          title = "Incidents by month",
          xaxis = list( title = "<b>Month</b>",
                        ticks = "inside",
                        tickangle = 15 ),
          yaxis = list( title = "<b>Incidents per month in NYC (2016)</b>",
                        ticks = "inside",
                        gridwidth = 2,
                        gridcolor = "lightgray" ) )
```


## Incidents by Borough

This bar plot shows the incidents per Borough. 

```{r plot-bars, echo=FALSE, width="100%"}
# Incident aggregation
incidents_by_city <- as.data.frame( table(parties$Borough) )
colnames( incidents_by_city ) <- c( "Borough", "IncidentCount" )
incidents_by_city$Borough <- factor(
  incidents_by_city$Borough,
  levels = unique(incidents_by_city$Borough)[order(
    incidents_by_city$IncidentCount, decreasing = TRUE )])

# Building the plot
plot_ly( x = ~incidents_by_city$Borough,
         y = ~incidents_by_city$IncidentCount,
         type = "bar",
               marker = list( size = 15, color = "#52006d" ),
               showlegend = FALSE,
               hoverinfo = "text",
               text = paste0( "<br><b>", incidents_by_city$Borough,
                              "</b></br> Incidents: ",
                              incidents_by_city$IncidentCount ) ) %>%
  layout( showlegend = FALSE,
          title = "Incidents by Borough",
          xaxis = list( title = "",
                        ticks = "inside",
                        tickangle = 90 ),
          yaxis = list( title = "<b>Incident count</b>",
                        ticks = "inside",
                        gridwidth = 2,
                        gridcolor = "lightgray" ),
          margin = list( r = 0,
                         b = 150 ) )
```

## Incidents by location type

Pie chart with aggregated incidents by location type.

```{r plot-pie, echo=FALSE, width="100%"}
# Incident aggregation
incidents_by_place <- as.data.frame( table(parties$Location.Type) )
colnames( incidents_by_place ) <- c( "Place", "IncidentCount" )
incidents_by_place$Place <- factor(
    incidents_by_place$Place,
    levels = unique(incidents_by_place$Place)[order(
         incidents_by_place$IncidentCount, decreasing = TRUE )] )

# Building the plot
plot_ly( labels = ~incidents_by_place$Place,
         values = ~incidents_by_place$IncidentCount,
         type = "pie",
         showlegend = FALSE,
         hoverinfo = "text",
         text = paste0( "<br><b>", incidents_by_place$Place,
                        "</b></br> Incidents: ",
                         incidents_by_place$IncidentCount ) ) %>%
  layout( showlegend = FALSE,
          title = "Incidents by Place (location type)" )
```

