---
title: "Cascade Volcanoes"
author: "Tinguaro Barreno"
date: "October 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This map shows the 82
[Cascade Volcanoes](https://en.wikipedia.org/wiki/List_of_Cascade_volcanoes),
formed as a result of subduction along the 
[Cascadia subduction zone](https://en.wikipedia.org/wiki/List_of_Cascade_volcanoes)
in the Pacific Northwest of North America.

```{r data-loading, echo=FALSE}
# CSV read
cascade_volcanoes <- read.csv("Cascade_Volcanoes.csv")
```
```{r the-plot, echo=FALSE}
# Library load
library("leaflet")

# This funcion sets the icon color by type inside the 'awesomeIcons'
getColor <- function( vei ) {
  sapply(cascade_volcanoes$LastEruptionVEI, function(vei) {
  if( vei == "1" || vei == "1?" ) {
    "green"
  } else if( vei == "2" || vei == "2?") {
    "lightgreen"
  } else if( vei == "3" || vei == "3?") {
    "orange"
  } else if( vei == "4" || vei == "4?") {
    "red"
  } else if( vei == "5" || vei == "5?") {
    "purple"
  } else {
    "lightgray"
  } })
}

# This function sets the icon text
getName <- function( vei )
{
  sapply(cascade_volcanoes$LastEruptionVEI, function(vei) {
    if( vei == "Unknown") {
      "-"
    } else {
      paste( "<b>", vei, "</b>" )
    }
  })
}

# Creates the icons
icons <- awesomeIcons(
  text = getName( cascade_volcanoes ),
  iconColor = "black",
  markerColor = getColor( cascade_volcanoes )
)

# Generate the interactive map:
cv_plot <- leaflet( cascade_volcanoes, width = "100%" ) %>%
  addProviderTiles( providers$Esri.NatGeoWorldMap ) %>%
  setView(lng = -121.49, lat = 46.206, zoom = 9) %>%
  addAwesomeMarkers( lng   = ~Longitude,
              lat   = ~Latitude,
              # color = ~pal(Type),
              icon  = icons,
              popup = ~paste("<h4>", Name, "</h4>",
                             "<br/>Type: <b>", Type, "</b><br/>",
                             "Elevation: ", ElevationFeets, "feets (",
                             ElevationMeters, "m)<br/>",
                             "Last eruption date: ", LastEruptionDate, "<br/>",
                             "Last eruption VEI: <tt>",
                             LastEruptionVEI, "</tt>") ) %>%
  addLegend( position = 'topright',
              title = "Last eruption VEI",
              colors = c("#71AF26", "#BBF970", "#F69730", "#D53E2A", "#CB4FB3", "#B0B0B0"), 
              labels = c("1", "2", "3", "4", "5+", "Unknown"),
              opacity = 0.7 ) %>%
  addMiniMap( position = 'bottomright', toggleDisplay = TRUE )

# Print it!
cv_plot
```



This document corresponds with the interactive **Leaflet** map required at the
peer-graded assignment for the "*Developing Data Products*" course on
[Coursera](https://www.coursera.org/learn/data-products/).

A desktop wallpaper image of the
[Mount Adams](https://en.wikipedia.org/wiki/Mount_Adams_(Washington))
was the inspiration for this project. I picked up the
[Wikipedia's list of Cascade Volcanoes](https://en.wikipedia.org/wiki/List_of_Cascade_volcanoes)
(with location coordinates) and convert the data to a CSV format
suitable for loading from **R**.

## Building the map

### Getting the data

The data is loaded directly from the CSV file (available in the
same **GitHub** repository):

```{r ref.label="data-loading", eval=FALSE}
```

### Explanatory Data Analysis

The volcanoes data consists in 10 variables. We will use
the `Latitude` and `Longitude` for the position markers, the
[Volcanic Explosivity Index](https://en.wikipedia.org/wiki/VEI)
for the markers color and text. The rest of the
information will be shown in the *popups*.

```{r eda}
str(cascade_volcanoes)
```


### The map

We want to show the volcanoes position over a topography map,
so we choose the "[Esri.NatGeoWorldMap](https://www.arcgis.com/home/item.html?id=b9b1b422198944fbbd5250b3241691b6)".
The map will be centered at the **Mount Adams**.

Each marker will have the VEI number if it's available and have
a color associated with this VEI (a legend is shown in the corner).
As seen in the [Leaflet documentation](https://rstudio.github.io/leaflet/markers.html),
we will use auxiliary functions to configure the color and labels for the
[AwesomeIcons](https://www.rdocumentation.org/packages/leaflet/versions/1.1.0/topics/awesomeIcons)
markers.

We will use *popups* to show aditional information about the volcanoes:
name, type, elevation, last eruption date and VEI.

```{r ref.label="the-plot", eval=FALSE}
```
